Supporting Information

### 2018-03-28 SupportingInformation.Rmd
gone through aphid count model selection section, editing

```{r read_aphid_data, echo=F}
# Aphid raw data for count
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

@@

# Aphid raw data for flight duration   
#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))

```

```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))}
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
#cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
```

@@

```{r tab_count_bestcop, echo=F, results='markup'}
#library(dplyr)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
             format='pandoc',
             caption = "Aphid_count : best fit copula family for each selective location pair\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='pandoc',
             caption = "Aphid_count : AICw for best fit copula family for each selective location pair\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_CvM,
             format='pandoc',
             caption = "Aphid_count : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_count_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_count_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_KS,
             format='pandoc',
             caption = "Aphid_count : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_count_p_KS}",
             booktabs=TRUE
             )
```

```{r tab_count_AICw_normal, echo=F, results='markup'}
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='pandoc',
             caption = "Aphid_count : AICw for 'Normal' copula family for each selective location pair\\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='pandoc',
             caption = "Aphid_count : model averaged lower tail dependence statistic for each selective location pair\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='pandoc',
             caption = "Aphid_count : model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Aphid_count : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```

Table \ref{tab_count_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data. Table \ref{tab_count_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_count_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_count_p_KS}] of the goodness of fit with atleast
`r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_count_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_count_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_count_LTdep_AICw} and Table \ref{tab_count_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_count_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.

The next section looks at tail dependence in a non-parametric way.

>>>>

<!--Load the raw data-->
```{r read_aphid_data, echo=F}
# Aphid raw data for count
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

@@

# Aphid raw data for flight duration   
#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
```

<!--Do the calculations-->
```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))}
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
#cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
```

As described in Methods, model selection among copulas was done separately for each pair of the 10 
sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
corresponding to locations. Table \ref{tab_count_bestcop} shows that, for the large majority of location 
pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{tab_count_AICw}) 
to the AIC weight 
of the normal copula family for the same pair (Table \ref{tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_count_p_KS}),
demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

```{r tab_count_bestcop, echo=F, results='markup'}
#library(dplyr)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
             format='pandoc',
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas. Abbreviations were given in Methods. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='pandoc',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for abundances of the green spruce aphid.\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_AICw_normal, echo=F, results='markup'}
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='pandoc',
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for abundances of the green spruce aphid, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_count_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_CvM,
             format='pandoc',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid. Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_count_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_KS,
             format='pandoc',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid. Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_KS}",
             booktabs=TRUE
             )
```

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{tab_count_LTdep_AICw} and \ref{tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.

```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='pandoc',
             caption = "Model-averaged lower tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='pandoc',
             caption = "Model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.

```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 5,
             format='pandoc',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```
